The Sub-Millimetre User Reduction Facility: an iterative map-maker for \scuba

### 2012-08-14 smurf_paper.tex
Fix missing label for appendix
Update references
SPIE references were missing page numbers and editor information.
Added some references to SURF and JSA into the text. Updated
a astro-ph reference that was actually published in 2009.

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{holland2006} is a new instrument for the 15-m James
Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
simultaneously images the sky in two broad bands centered over 450 and
850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
half-maximum (FWHM) point spread functions (PSFs). The focal planes at
each wavelength are populated with 4 rectangular subarrays, consisting
of $40 \times 32$ bolometers each, and subtend a nearly contiguous $7'
\times 7'$ footprint on the sky.  Ultimately, the instrument will also
be equipped with a Fourier transform spectrometer
\citep[FTS-2,][]{gom2010}, and a polarimeter
\cite[POL-2,][]{bastien2005}. This paper describes the properties of
\scuba\ data that are relevant to producing maps of the imaging data,
and the Submillimetre User Reduction Facility, SMURF, a software
package for perfoming the reduction. The analysis of FTS-2 and POL-2
data will be described at a later date once these additional
instruments have been commissioned. The details of the instrument
design, performance, and its calibration are given in two companion
papers: \citet{holland2012} and \citet{dempsey2012}.

@@

There are three general styles of map-making that are relevant to
reducing bolomter data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, e.g., using baseline removal,
filters, and then simply re-gridding the data into a map. Such was the
basic recipe for the baseline reduction of chopped data from SCUBA-2's
predecessor SCUBA (reference?), and MAMBO, another camera from the
same generation. Generally speaking, such methods are fast, although
dependending on the science goals and noise properties of the data,
they may not achieve the best theoretical noise performance on the
angular scales of interest. A particular case of this method was
successfully used for the reduction of data from BOLOCAM
\citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.][]{scott2008}, using ``principal component analysis''
(PCA) cleaning. A statistical black-box is used to remove the most
correlated components of the bolometer signals, enabling the detection
of point-sources very close to the theoretical white-noise limits of
the detectors with reasonable computation times. However, PCA cleaning
is not a good generic solution for producing maps of extended
structures, since large astronomical sources produce correlated
signals amongst many detectors, and are removed by this procedure.

@@

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromosing structures on angular scales larger than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data faster than they are taken, on single high-end
desktop computers, and it has successfully been used as part of a
real-time pipeline offering feedback to observers at the
telescope. The pipeline is also used to generate near-science grade
products at the Canadian Astronomical Data Centre (CADC). Details of
the pipeline design are given in \citet{gibb2005}

@@

Have real astronomical sources been detected using the matched-filter?
For both the smoothed signal and jackknife maps, orange circles denote
3.8-$\sigma$ peaks. While not justified here, this threshold is fairly
typical for other ground-based submillimetre surveys in recent years
\citep[e.g.,][]{coppin2006,perera2008,weiss2009} leading to estimated
false-identification rates of order $\sim$5\%, and is chosen here as a
convenient reference. In the former, 10 peaks are found, whereas there
is only 1 in the latter. However, this test does not preclude the
possibility that some correlated noise made it into the jackknife
map, in which case the estimated SNRs are misleading.

One simple test of the calculated noise properties is to compare the
signal and jackknife SNR distributions with ideal Gaussians. The top
panel of Fig.~\ref{fig:lockman_hist} shows the whitened (but not match
filtered) signal (blue) and jackknife (histograms), along with a
Gaussian (mean 0, standard deviation 1, and area normalized to the
number of map pixels) as a dashed line. In this case, it is clear that
the SNR distributions for both maps are nearly indistinguishable from
the theoretical distribution of white noise. This result shows us
that: (i) the whitening filter appears to have removed correlated
large-scale noise since the jackknife map histogram is consistent with
white noise; and (ii) any potential astronomical signals are small
compared to the typical white noise in most map pixels (unsurprising
given the appearance of Fig.~\ref{fig:lockman_maps}b). Next, we
examine the SNR histograms for maps processed with the matched filter
in the bottom panel of Fig.~\ref{fig:lockman_hist}. Again, the
histogram of the jackknife SNR data appears consistent with pure
noise. However, the signal map now deviates significantly from a
Gaussian distribution, with a clear \emph{positive} tail (as one would
expect for emitting sources). In fact, integrating the positive tails
beyond our 3.8-$\sigma$ source-detection threshold (vertical solid
line) yields 229 map pixels in the signal map, compared with 3 in the
jackknife map (out of 80603 pixels in the entire region). Note that
given the small pixel size of the map, several pixels generally
contribute to each peak. Note that this procedure gives only a flavour
of the analysis that is usually required to produce a robust source
lists. Additional tests, along with a careful consideration of
completeness and false-positive rates usual require a series of Monte
Carlo simulations that are beyond the scope of this work. We direct
the interested reader to a selection of papers on the subject:
\citet{scott2002,coppin2006,perera2008,weiss2009} and
\citet{chapin2011}.

@@

%------------------------------------------------------------------------------
\appendix
\section[]{Step correction in detail}

>>>>

The Submillimetre Common User Bolometer Array 2
\citep[\scuba,][]{2006SPIE.6275E..45H,holland2012} is a new instrument for the 15-m James
Clerk Maxwell Telescope (JCMT) on Mauna Kea, Hawai'i. The camera
simultaneously images the sky in two broad bands centered over 450 and
850\,\micron, with approximately 7.5 and 14.5\,arcsec full-width at
half-maximum (FWHM) point spread functions (PSFs). The focal planes at
each wavelength are populated with 4 rectangular subarrays, consisting
of $40 \times 32$ bolometers each, and subtend a nearly contiguous $7'
\times 7'$ footprint on the sky.  Ultimately, the instrument will also
be equipped with a Fourier transform spectrometer
\citep[FTS-2,][]{gom2010}, and a polarimeter
\cite[POL-2,][]{2011ASPC..449...68B}. This paper describes the properties of
\scuba\ data that are relevant to producing maps of the imaging data,
and the Submillimetre User Reduction Facility, SMURF, a software
package for perfoming the reduction written using the Starlink
Software Environment \citep{2009ASPC..411..418J}. The analysis
of FTS-2 and POL-2 data will be described at a later date once these additional
instruments have been commissioned. The details of the instrument
design, performance, and its calibration are given in two companion
papers: \citet{holland2012} and \citet{dempsey2012}.

@@

There are three general styles of map-making that are relevant to
reducing bolomter data in the literature. The simplest ``direct
methods'' involve some basic processing of the data to remove as much
noise contamination as possible, e.g., using baseline removal,
filters, and then simply re-gridding the data into a map. Such was the
basic recipe for the baseline reduction of chopped data from SCUBA-2's
predecessor SCUBA \citep{1998ASPCd..145..216J,2000ASPC..216..559J},
and MAMBO, another camera from the same generation. Generally
speaking, such methods are fast, although dependending on the science
goals and noise properties of the data,
they may not achieve the best theoretical noise performance on the
angular scales of interest. A particular case of this method was
successfully used for the reduction of data from BOLOCAM
\citep[e.g.][]{laurent2005}, and its younger sibling AzTEC
\citep[e.g.][]{scott2008}, using ``principal component analysis''
(PCA) cleaning. A statistical black-box is used to remove the most
correlated components of the bolometer signals, enabling the detection
of point-sources very close to the theoretical white-noise limits of
the detectors with reasonable computation times. However, PCA cleaning
is not a good generic solution for producing maps of extended
structures, since large astronomical sources produce correlated
signals amongst many detectors, and are removed by this procedure.

@@

A reasonable model for correlated noise in \scuba\ is a single
``common-mode'' signal seen by all of the bolometers. Iterative
estimation and removal of this signal significantly lowers the $1/f$
knee, without compromosing structures on angular scales larger than
the array footprint. Residual independent drifts at lower frequencies
are removed with an iterative high-pass filter. This strategy enables
SMURF to reduce data faster than they are taken, on single high-end
desktop computers, and it has successfully been used as part of a
real-time pipeline offering feedback to observers at the
telescope. The pipeline is also used to generate near-science grade
products using the JCMT Science Archive \citep{2011ASPC..442..203E}
hosted by the Canadian Astronomical Data Centre (CADC). Details of
the pipeline design are given in \citet{gibb2005} and
\citet{2008AN....329..295C}.

@@

Have real astronomical sources been detected using the matched-filter?
For both the smoothed signal and jackknife maps, orange circles denote
3.8-$\sigma$ peaks. While not justified here, this threshold is fairly
typical for other ground-based submillimetre surveys in recent years
\citep[e.g.,][]{coppin2006,perera2008,2009ApJ...707.1201W} leading to estimated
false-identification rates of order $\sim$5\%, and is chosen here as a
convenient reference. In the former, 10 peaks are found, whereas there
is only 1 in the latter. However, this test does not preclude the
possibility that some correlated noise made it into the jackknife
map, in which case the estimated SNRs are misleading.

One simple test of the calculated noise properties is to compare the
signal and jackknife SNR distributions with ideal Gaussians. The top
panel of Fig.~\ref{fig:lockman_hist} shows the whitened (but not match
filtered) signal (blue) and jackknife (histograms), along with a
Gaussian (mean 0, standard deviation 1, and area normalized to the
number of map pixels) as a dashed line. In this case, it is clear that
the SNR distributions for both maps are nearly indistinguishable from
the theoretical distribution of white noise. This result shows us
that: (i) the whitening filter appears to have removed correlated
large-scale noise since the jackknife map histogram is consistent with
white noise; and (ii) any potential astronomical signals are small
compared to the typical white noise in most map pixels (unsurprising
given the appearance of Fig.~\ref{fig:lockman_maps}b). Next, we
examine the SNR histograms for maps processed with the matched filter
in the bottom panel of Fig.~\ref{fig:lockman_hist}. Again, the
histogram of the jackknife SNR data appears consistent with pure
noise. However, the signal map now deviates significantly from a
Gaussian distribution, with a clear \emph{positive} tail (as one would
expect for emitting sources). In fact, integrating the positive tails
beyond our 3.8-$\sigma$ source-detection threshold (vertical solid
line) yields 229 map pixels in the signal map, compared with 3 in the
jackknife map (out of 80603 pixels in the entire region). Note that
given the small pixel size of the map, several pixels generally
contribute to each peak. Note that this procedure gives only a flavour
of the analysis that is usually required to produce a robust source
lists. Additional tests, along with a careful consideration of
completeness and false-positive rates usual require a series of Monte
Carlo simulations that are beyond the scope of this work. We direct
the interested reader to a selection of papers on the subject:
\citet{scott2002,coppin2006,perera2008,2009ApJ...707.1201W} and
\citet{chapin2011}.

@@

%------------------------------------------------------------------------------
\appendix
\section[]{Step correction in detail}
\label{appendix:steps}
